Categorical Predictor Embeddings in Actuarial Modeling

### 2021-03-30 manuscript/manuscript.Rmd
Clean copy in first sections

This approach to modelling categorical data has several advantages over more traditional treatments of categorical data.
Using categorical embeddings instead of traditional techniques has been shown to increase predictive accuracy of 
models, for example, see [@Richman2018] in the context of pricing. Models incorporating categorical embeddings can be
pre-calibrated to traditional actuarial models, increasing the speed with which these models can be calibrated and 
leading to models with better explainability [@wuthrich2019yes]. Finally, the similarity between the vectors learned 
for different categories can be inspected, sometimes leading to insights into the workings of models, see, for example,
[@Kuo2019].

@@

The rest of this manuscript is organized as follows. Section \ref{lit_review} reviews recent applications of embeddings
in the actuarial literature. Section \ref{defns} provides the notation used in the paper and defines GLMs, neural 
networks and related modelling concepts, including embeddings and attention. In Section \ref{modelling}, we provide 
initial models for the NFIP dataset and consider the influence of hyperparameter choices on the results of the neural 
network model. Section \ref{tl} considers how successfully the embeddings used in the neural network model can be 
transferred to a GLM model of the same data. Extensions to the modelling approach are considered in Section \ref{extend} 
and we focus on attention based models in Section \ref{attention}. The interpretability of embedding layers is addressed
in Section \ref{interpretation}. Finally, Section \ref{Conclusions} provides a discussion of the results of this paper
and considers avenues for future research.

@@

Categorical data are usually modelled within GLMs and other predictive models using indicator variables which capture 
the effect of each level of the category, see, for example, Section 2 in [@Goldburd], using one of two main encoding 
schemes: dummy-coding and one-hot encoding. Dummy-coding, used in the popular \texttt{R} statistical software, assigns 
one level of the category as a baseline, for which an indicator variable is not calibrated, and the rest of the levels 
are assigned indicator variables, thus, producing estimates within the model of how the effects of each level differ 
from the baselines. One-hot encoding, often used in machine learning, is similar to dummy-coding, but assigns 
indicator variables to each level, in other words, calibrates an extra indicator variable compared with dummy-coding.

A different approach to modelling categorical data is credibility theory (see [@Buhlmann2006] for an overview), which, 
in the context of rating, can be applied to derive premiums that reflect the experience of a particular policyholder, by 
estimating premiums as a weighted average between the premium produced using the collective experience (i.e. of all 
policyholders) and the premium produced using the experience of the particular policyholder. The weight used in this 
average is called a credibility factor and is calculated with reference to te variability of the policyholder experience 
relative to the variability of the group experience. In this context, the implicit categorical variable is the 
policyholder under consideration.

@@

Embedding layers represent a different approach to the problem of modelling categorical data that was recently 
introduced in an actuarial context. Note that in the next section, we reflect on similarities between the conventional 
approaches discussed above and embedding layers. [@Richman2018] reviewed the concept of embedding layers and connected 
the sharing of information across categories to the familiar concept of credibility theory. In that work, two 
applications of embedding layers were demonstrated. The first of these was in a Property and Casualty (P&C) pricing 
context, it was shown that the out-of-sample accuracy of a neural network trained to predict claims frequencies on motor 
third party liability was enhanced by modelling the categorical variables within this dataset using embedding layers. 
Second, a neural network with embedding layers was used to model all of the mortality rates in the Human Mortality 
Database, where the differences in population mortality across countries and the differences in mortality at different 
ages were modelled with embedding layers, again producing more accurate out of sample performance than the other models 
tested.

Contemporaneous with that work is the DeepTriangle model of [@Kuo2019], which applied recurrent neural networks to the 
problem of Incurred but not Reported (IBNR) loss reserving, to model jointly the paid and incurred losses in the 
Schedule P dataset. Embedding layers were used to capture the effect of differences in reserving delays and loss ratios 
for each company in the Schedule P dataset. Evaluating the results of the DeepTriangle method showed that the out of 
sample performance of the model (tested against the lower triangles in the Schedule P dataset) exceeded that of 
traditional IBNR reserving techniques.

Many other applications of embeddings have subsequently appeared in the actuarial literature. Within mortality 
forecasting, [@Richman2019d] and [@Richman2020] both apply embeddings layers to model and forecast mortality rates on a 
large scale. [@wuthrich2019yes] discussed how embeddings can be calibrated using GLM techniques and then incorporated 
into a combined actuarial neural network, with subsequent contributions in P&C pricing by [@schelldorfer2019nesting] and 
in IBNR reserving by [@Gabrielli2019c] and [@Gabrielli2019]. Other applications in IBNR reserving are in [@Kuo2020] and 
[@Delong2020a] who use embedding layers to model individual claims development.

@@

A categorical variable $X_j, j \in J_{cat}$ takes as its value only one of a finite number of labels. Let the set of 
labels be $\mathcal{P}^j = \{p^j_1, p^j_2,\dots, p^j_{n_{\mathcal{P}^j}} \}$, where $n_\mathcal{P}^j = |\mathcal{P}^j|$ 
is the cardinality or number of levels, in $\mathcal{P}^j$. One-hot encoding maps each value $x_{i,j}$ of $X_j$ to 
$n_\mathcal{P}^j$ indicator variables, which take a value of $1$ if the label of $x_{i,j}$ corresponds to the level of 
the indicator variable, and $0$ otherwise. An example of one-hot encoding is shown in Table \ref{onehot}.

@@

One-hot encoding is often used in the machine learning community while the statistical community often favors dummy 
coding, which, instead of assigning $n_p^j$ indicator variables, assigns one of the levels of the categories as a 
baseline, and maps all of the other $n_p^j -1$ variables to indicator variables. An example of dummy encoding is shown 
in Table \ref{dummy}.

@@

These coefficients represent the effect that each level of the categorical variable will have on the outcome. In the 
case that there are no other variables available in the dataset, then the coefficients will reflect the average value of 
the outcomes for that level of the categorical variable. For example, suppose that the categorical variable is a 
policyholder identifier, and the outcomes are the value of claims in different years, then the coefficients will reflect 
the average annual claims for each policyholder based on the experience. In other words, both of these encoding schemes 
give full credibility to the data available for each category, thus, even if a relatively small amount of data is 
available for a specific policyholder, the coefficient that is calibrated will only reflect that data. On the other 
hand, a foundational technique within actuarial work is the application of credibility methods, which are used for 
experience rating and other applications. These techniques provide an estimate that reflects not only the experience of 
the individual policyholder but also that of the collective, based on an estimate of how credible the data for each 
individual is. While we have described the application of credibility in a simple univariate context, it is also 
possible to apply credibility considerations within GLMs. using models known as Generalized Linear Mixed Models or 
GLMMs, and we refer to [@Klinker2010] for more details.

@@

Neural networks are flexible machine learning models that have recently been applied to a number of problems with 
Property and Casualty (P&C) insurance. Here, we provide a brief overview of these models, and refer the reader to 
[@Richman2018] for a more detailed overview. Neural networks are characterized by multiple layers of non-linear 
regression functions that are used to learn a new representation of the data input to the network that is then used to 
make predictions. Here we focus on the most common type of neural networks, which are fully connected networks (FCNs), 
which provide as the output of each set of non-linear functions to the subsequent layer of functions. Formally, a 
$K$-layer neural network is:

@@

where the regression parameters (weights) for each layer $$k \in [1;K]$$ are represented by the matrices $a_k$ and the 
intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear 
regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the tanh function or 
the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') as follows. First a 
loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ and the 
predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of 
the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation).
Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

If $K$ is set equal to $1$, then Equation \ref{NN} reduced to nothing more than a GLM. A neural network with $K=2$ is 
called a shallow neural network and for $K \geq 2$, the network is called a deep neural network. The matrix $X$ of data 
input to the network can be composed of both continuous variables as well as categorical variables, which can be 
pre-processed using one-hot or dummy encoding. As mentioned above, a different option is to use encodings, which we 
discuss in more detail next.

@@

Common issues with the traditional encoding schemes for categorical data occur when the number of levels for each 
variable is very large. Often, in these cases, models do not converge quickly, and the very large matrices that result 
from applying these schemes often cause computational difficulties. Besides for these practical issues, a deeper issue 
is that one-hot or dummy encoded data assumes that each category is entirely independent of the rest of the categories, 
in other words, there are no similarities between categories that could enable more robust estimation of models. In 
technical terms, this is because the columns of the matrices created by one-hot encoding are all orthogonal to each 
other. (These arguments appear in a similar form in [@Guo2016].) Solutions to these problems are provided by embedding 
layers.

@@

$$
z_\mathcal{P^j} : \mathcal{P^j} \to \mathbb{R}^{q_\mathcal{P^j} }, \quad \quad p^j \mapsto  z_\mathcal{P^j}(p),
$$

where $q_\mathcal{P^j}$ is the dimension of the embedding for the $j$th categorical variable and $z_\mathcal{P^j}(.)$ is 
a function that maps from the particular element of the labels $p$ to the embedding space. Equation \ref{eq:embedding} 
states that an embedding maps a level of a categrical variable to a numerical vector. This function is left implicit, 
meaning to say, we allow the embeddings to be derived during the process of fitting the model and do not attempt to 
specify exactly how the embeddings can be derived from the input data. In Table \ref{embed} we show an example of two 
dimensional embeddings for the state variables, where these have been generated randomly.

@@

data.frame(sample_dat, dimension1 = rnorm(5), dimension2 = rnorm(5)) %>%
  knitr::kable(label = "", caption = "Example (random) embeddings of the state variable")
```

@@

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], and, more recently, has been applied to computer vision and tabular modelling tasks. Attention mechanisms allow deep neural nets to weight the covariates entering the model in a flexible manner; these types of applications in natural langauge processing are demonstrated in, for example, [@Vaswani2017a]. In an actuarial context, attention mechanisms can be understood as giving greater weight to covariates that are important for the modelling task. In this section we provide a theoretical introduction to attention mechanisms and then discuss applications of attention to modeling tabular data. In the following sections of this paper, we then consider whether adding attention to an model that uses embeddings can provide better predictive accuracy. The following is intended as a high level review of some of the relevant theory behind attention based approaches; for more technical details the sources quoted in this section can be consulted.

@@

```{=latex}
\begin{equation}
A: \mathbb{R}^{(q \times d) \times (q \times d) \times (q \times d)} \to \mathbb{R}^(q \times d), ~~
(Q, K, V) ~\mapsto ~ V^* = A(Q, K, V),
\label{eq:attention}
\end{equation}
```
where $Q \in \mathbb{R}^{q \times d}$ is a matrix of query vectors, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors, $X' \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $X^* \in \mathbb{R}^{q \times d}$, the output of the attention mechanism, is a new matrix of values. Several different options for the mapping \eqref{eq:attention} exist in the literature; here we describe the self-attention mechanism used in [@Vaswani2017a]. In more detail, consider that $X'$ is a matrix of covariates relating to a regression problem, which, in the case of tabular data, will usually be composed of embeddings of the same dimension for several different categorical or numerical variables; in the case discussed here we have $q$ covariates in the form of embeddings of dimension $d$. For clarity, note that in Equation \ref{eq:embed_to_nn} we have used the symbol $X'$ to denote the vector of embeddings and numerical inputs which is passed into the neural network. In what follows, we now use the symbol $X'$ to refer to the matrix of embeddings of dimension $q \times d$.

We wish to apply the attention mechanism to $X'$ to assign weights to the most important embeddings for the regression problem; this is done from the \`\`perspective" of each covariate in $x'$. To do this, we formulate the query matrix $Q$ which contains information about the regression problem at hand; this is usually done by passing $X'$ through a FCN trained to derive queries from the covariates. Similarly, the matrix $K$ contains the relevance of each row of $V$ for the regression problem and is also usually derived by passing each row of $X'$ through another FCN. Finally, the matrix $X'$can be processed with a third FCN or left as is. The first step of the attention mechanism is to derive a matrix $A \in \mathbb{R}^{q \times q}$ of attention scores:

```{=latex}
\begin{equation}
\mathbb{R}^{(q \times d) \times (q \times d)} \to \mathbb{R}^{q \times q}, ~~
(Q, K) ~\mapsto ~ A = Q~K^T.
\label{eq:attention_scores}
\end{equation}
```
The attention scores $A$ are then normalized, or made to sum to unity, across the columns of $A$ using a so-called softmax function i.e.

@@

```{=latex}
\begin{equation}
\mathbb{R}^{(q \times q) \times (q \times d)} \to \mathbb{R}^{q \times d}, ~~
(A^*, x') ~\mapsto ~ X^* = A^*~X'.
\label{eq:attention_output}
\end{equation}
```
In most applications of attention, Equation \ref{eq:attention_scores} is modified by dividing the attention scores by $d^{\frac{1}{2}}$.

@@

Self-attention is the main building block of the current state of the art model for natural language processing, which is the Transformer model [@Vaswani2017]. A basic Transformer model applies self-attention to the covariate matrix, $X'$ which is fed into the model, to produce a new matrix of covariates, $X^*$. These two matrices, the input matrix $X'$ and the output matrix $X^*$ are added (in an element-wise manner) together and then normalized. This matrix is then further processed through a neural network and subsequently, these processed covariates are used in the machine learning task. Multiple Tranformer network "blocks" can be added to a network to create a deep Transformer network.

@@

```{=latex}
\begin{eqnarray}
  X^* = Attn(X') \\
  z^1 = \sigma(a_1.X^* + b_1),
\label{eq:embed_to_attn_to_nn}
\end{eqnarray}
```
where $X' \in \mathbb{R}^{q \times d}$ is the matrix of embeddings after concatenating $q$ embeddings of dimension $d$ together and $X^*$ represents the matrix of embeddings $X'$ after processing with attention. Below, we demonstrate the effect of adding this simple application of attention to a tabular model.

[@Huang2020] refines this approach in a model they call the TabTransformer, which implements two main changes to the approach shown in Equations \ref{eq:embed_to_attn_to_nn}. The first of these is that, instead of using the matrix of embeddings after processing with self-attention ($X'$) directly in the model, the output of a Transformer model is used instead. The second of these refinements is that for each of the $q$ covariates with an embedding in the matrix $X'$, we include a new row embedding which denotes which covariate each embedding relates to. Thus, in the TabTransformer model, we augment the embedding matrix with extra columns, so that the dimension of this matrix is $q \times (d + d_col)$, where $d_col$ is the dimension of the row embedding. [@Huang2020] find that including this identifier improves the performance of the model. We compare the results of the TabTransformer model to the other models tested below.

Finally, a different way of incorporating attention into the modeling of tabular data has been proposed by [@Ark2019], who refer to their model as TabNet. In summary, the main idea of the TabNet model is to try emulate the excellent performance of decision trees on tabular data modelling using deep neural networks. This is done by structuring the neural networks to emulate a key aspect of decision tree modelling, which is the efficient selection of relevant covariates for the model task. Within TabNet, an attention based model is used to select features by estimating attention scores for each covariate within $X$. The covariates are multiplied by the attention scores to downweight less relevant features for the prediction problem.

# Predictive Modeling with Embeddings

@@

# Extending the model {#extend}

Investigate approaches for using embedding layers for numeric variables, or numeric variables that were captured as 
categorical variables (Can the modelling of numerical variables benefit from the application of embedding layers?) How 
should this best be done? CatBoost - cut into 256 groups smoothness Map numeric to a dense layer and use that as 
embedding - basis expansion Add a distribution to the embeddings Can insights from credibility theory lead to enhanced embeddings? Take credibility mixture over embeddings Cluster embeddings PCA of embeddings

# Attention based modelling {#attn}

>>>>

This approach to modelling categorical data has several advantages over more traditional treatments of categorical data.
Using categorical embeddings instead of traditional techniques has been shown to increase predictive accuracy of 
models, for example, see @Richman2018 in the context of pricing. Models incorporating categorical embeddings can be
pre-calibrated to traditional actuarial models, increasing the speed with which these models can be calibrated and 
leading to models with better explainability [@wuthrich2019yes]. Finally, the similarity between the vectors learned 
for different categories can be inspected, sometimes leading to insights into the workings of models, see, for example,
@Kuo2019.

@@

The rest of this manuscript is organized as follows. Section \ref{lit_review} reviews recent applications of embeddings
in the actuarial literature. Section \ref{defns} provides the notation used in the paper and defines GLMs, neural 
networks and related modelling concepts, including embeddings and attention. In Section \ref{modelling}, we provide 
initial models for the NFIP dataset and consider the influence of hyperparameter choices on the results of the neural 
network model. That section also considers how successfully the embeddings used in the neural network model can be 
transferred to a GLM model of the same data and how learned embeddings can be interpreted. We focus on attention based models 
in Section \ref{attention_models}. Finally, Section \ref{Conclusions} provides a discussion of the results of this paper
and considers avenues for future research.

@@

Categorical data are usually modelled within GLMs and other predictive models using indicator variables which capture 
the effect of each level of the category, see, for example, Section 2 in @Goldburd, using one of two main encoding 
schemes: dummy-coding and one-hot encoding. Dummy-coding, used in the popular \texttt{R} statistical software, assigns 
one level of the category as a baseline, for which an indicator variable is not calibrated, and the rest of the levels 
are assigned indicator variables, thus, producing estimates within the model of how the effects of each level differ 
from the baselines. One-hot encoding, often used in machine learning, is similar to dummy-coding, but assigns 
indicator variables to each level, in other words, calibrates an extra indicator variable compared with dummy-coding.

A different approach to modelling categorical data is credibility theory (see @Buhlmann2006 for an overview), which, 
in the context of rating, can be applied to derive premiums that reflect the experience of a particular policyholder, by 
estimating premiums as a weighted average between the premium produced using the collective experience (i.e. of all 
policyholders) and the premium produced using the experience of the particular policyholder. The weight used in this 
average is called a credibility factor and is calculated with reference to te variability of the policyholder experience 
relative to the variability of the group experience. In this context, the implicit categorical variable is the 
policyholder under consideration.

@@

Embedding layers represent a different approach to the problem of modelling categorical data that was recently 
introduced in an actuarial context. Note that in the next section, we reflect on similarities between the conventional 
approaches discussed above and embedding layers. @Richman2018 reviewed the concept of embedding layers and connected 
the sharing of information across categories to the familiar concept of credibility theory. In that work, two 
applications of embedding layers were demonstrated. The first of these was in a Property and Casualty (P&C) pricing 
context, it was shown that the out-of-sample accuracy of a neural network trained to predict claims frequencies on motor 
third party liability was enhanced by modelling the categorical variables within this dataset using embedding layers. 
Second, a neural network with embedding layers was used to model all of the mortality rates in the Human Mortality 
Database, where the differences in population mortality across countries and the differences in mortality at different 
ages were modelled with embedding layers, again producing more accurate out of sample performance than the other models 
tested.

Contemporaneous with that work is the DeepTriangle model of @Kuo2019, which applied recurrent neural networks to the 
problem of Incurred but not Reported (IBNR) loss reserving, to model jointly the paid and incurred losses in the 
Schedule P dataset. Embedding layers were used to capture the effect of differences in reserving delays and loss ratios 
for each company in the Schedule P dataset. Evaluating the results of the DeepTriangle method showed that the out of 
sample performance of the model (tested against the lower triangles in the Schedule P dataset) exceeded that of 
traditional IBNR reserving techniques.

Many other applications of embeddings have subsequently appeared in the actuarial literature. Within mortality 
forecasting, @Richman2019d and @Richman2020 both apply embeddings layers to model and forecast mortality rates on a 
large scale. @wuthrich2019yes discussed how embeddings can be calibrated using GLM techniques and then incorporated 
into a combined actuarial neural network, with subsequent contributions in P&C pricing by @schelldorfer2019nesting and 
in IBNR reserving by @Gabrielli2019c and @Gabrielli2019. Other applications in IBNR reserving are in @Kuo2020 and 
@Delong2020a who use embedding layers to model individual claims development.

@@

A categorical variable $X_j, j \in J_{cat}$ takes as its value only one of a finite number of labels. Let the set of 
labels be $\mathcal{P}^j = \{p^j_1, p^j_2,\dots, p^j_{n_{\mathcal{P}^j}} \}$, where $n_\mathcal{P}^j = |\mathcal{P}^j|$ 
is the cardinality or number of levels, in $\mathcal{P}^j$. One-hot encoding maps each value $x_{i,j}$ of $X_j$ to 
$n_\mathcal{P}^j$ indicator variables, which take a value of $1$ if the label of $x_{i,j}$ corresponds to the level of 
the indicator variable, and $0$ otherwise. An example of one-hot encoding is shown in Table \@ref(tab:onehot).

@@

One-hot encoding is often used in the machine learning community while the statistical community often favors dummy 
coding, which, instead of assigning $n_p^j$ indicator variables, assigns one of the levels of the categories as a 
baseline, and maps all of the other $n_p^j -1$ variables to indicator variables. An example of dummy encoding is shown 
in Table \@ref(tab:dummy).

@@

These coefficients represent the effect that each level of the categorical variable will have on the outcome. In the 
case that there are no other variables available in the dataset, then the coefficients will reflect the average value of 
the outcomes for that level of the categorical variable. For example, suppose that the categorical variable is a 
policyholder identifier, and the outcomes are the value of claims in different years, then the coefficients will reflect 
the average annual claims for each policyholder based on the experience. In other words, both of these encoding schemes 
give full credibility to the data available for each category, thus, even if a relatively small amount of data is 
available for a specific policyholder, the coefficient that is calibrated will only reflect that data. On the other 
hand, a foundational technique within actuarial work is the application of credibility methods, which are used for 
experience rating and other applications. These techniques provide an estimate that reflects not only the experience of 
the individual policyholder but also that of the collective, based on an estimate of how credible the data for each 
individual is. While we have described the application of credibility in a simple univariate context, it is also 
possible to apply credibility considerations within GLMs. using models known as Generalized Linear Mixed Models or 
GLMMs, and we refer to @Klinker2010 for more details.

@@

Neural networks are flexible machine learning models that have recently been applied to a number of problems with 
Property and Casualty (P&C) insurance. Here, we provide a brief overview of these models, and refer the reader to 
@Richman2018 for a more detailed overview. Neural networks are characterized by multiple layers of non-linear 
regression functions that are used to learn a new representation of the data input to the network that is then used to 
make predictions. Here we focus on the most common type of neural networks, which are fully connected networks (FCNs), 
which provide as the output of each set of non-linear functions to the subsequent layer of functions. Formally, a 
$K$-layer neural network is:

@@

where the regression parameters (weights) for each layer $k \in [1;K+1]$ are represented by the matrices $a_k$ and the 
intercept terms are represented by $b_k$. Whereas the calculation inside each of the layers is nothing more than linear 
regression, $\sigma$ represents the non-linear part of each layer. Choices for $\sigma$ are often the hyperbolic tangent 
(tanh) function or the rectified linear unit (ReLu) $max(0,x)$. The parameters of the network are estimated ('trained') 
as follows. First a loss function $L(.,.)$ is specified for the network that measures the difference between the observed data $y$ 
and the predictions of the network $\hat{y}$, for example, the Mean Squared Error ($(y - \hat{y})^2$). Then, the parameters of 
the network are changed such that the loss decreases (formally, this is done using the technique of backpropagation).
Finally, training is stopped once the predictive performance of the network on unseen data is suitably good.

If $K$ is set equal to $1$, then Equation \ref{NN} reduces to nothing more than a GLM. A neural network with $K=2$ is 
called a shallow neural network and for $K \geq 2$, the network is called a deep neural network. The matrix $X$ of data 
input to the network can be composed of both continuous variables as well as categorical variables, which can be 
pre-processed using one-hot or dummy encoding. As mentioned above, a different option is to use encodings, which we 
discuss in more detail next.

@@

Common issues with the traditional encoding schemes for categorical data occur when the number of levels for each 
variable is very large. Often, in these cases, models do not converge quickly, and the very large matrices that result 
from applying these schemes often cause computational difficulties. Besides for these practical issues, a deeper issue 
is that one-hot or dummy encoded data assumes that each category is entirely independent of the rest of the categories, 
in other words, there are no similarities between categories that could enable more robust estimation of models. In 
technical terms, this is because the columns of the matrices created by one-hot encoding are all orthogonal to each 
other. (These arguments appear in a similar form in @Guo2016.) Solutions to these problems are provided by embedding 
layers.

@@

\begin{equation} 
\label{embed_map}
  z_\mathcal{P}^j : \mathcal{P}^j \to \mathbb{R}^{q_\mathcal{P}^j }, \quad \quad p^j \mapsto  z_\mathcal{P}^j(p),
\end{equation}

where $q_\mathcal{P}^j$ is the dimension of the embedding for the $j$th categorical variable and $z_\mathcal{P}^j(.)$ is 
an implicit function that maps from the particular element of the labels $p$ to the embedding space. Equation \ref{embed_map} 
states that an embedding maps a level of a categorical variable to a numerical vector. This function is left implicit, 
meaning to say, we allow the embeddings to be derived during the process of fitting the model and do not attempt to 
specify exactly how the embeddings can be derived from the input data. In Table \@ref(tab:embed) we show an example of two 
dimensional embeddings for the state variables, where these have been generated randomly.

@@

data.frame(sample_dat, dimension1 = rnorm(5), dimension2 = rnorm(5)) %>%
  knitr::kable(label = "embed", caption = "Example (random) embeddings of the state variable")
```

@@

Attention mechanisms have been widely applied within the deep learning literature to give more flexibility to RNNs [@Bahdanau2015a] 
and subsequently as a replacement for FCNs in the so-called Transformer architecture, which is now widely used in natural language processing [@Vaswani2017a], 
and, more recently, has been applied to computer vision and tabular modelling tasks. 
Attention mechanisms allow deep neural nets to weight the covariates entering the model in a flexible manner; these types of applications in natural language 
processing are demonstrated in,
for example, @Vaswani2017a. In an actuarial context, attention mechanisms can be understood as giving greater weight to covariates that are important for the modelling task. 
In this section we provide a theoretical introduction to attention mechanisms and then discuss applications of attention to modeling tabular data. 
In the following sections of this paper, we then consider whether adding attention to an model that uses embeddings can provide better predictive accuracy. 
The following is intended as a high level review of some of the relevant theory behind attention based approaches; for more technical details the sources quoted in this section can be consulted.

At a high level, attention mechanisms reweight the covariates $X$ within a predictive model to give greater weight to covariates that are more
predictive for a particular problem. The attention weights can be derived in several different manners. Earlier works, such as @Bahdanau2015a
used so-called "additive" attention scores derived as the output of a neural network taking the covariates as inputs. Currently, the most popular
form of attention is the "scaled dot-product" attention of @Vaswani2017a which we describe in what follows. We also mention that usually, so-called
"self-attention" is applied, meaning to say that the importance scores for the covariates are derived using the covariates themselves (i.e. in what follows,
the attention scores are derived in Equation \ref{eq:attention_scores} using the transformed covariates). Other options, such as using a different source of
data to derive the attention scores, are also possible.

@@

```{=latex}
\begin{equation}
X^*: \mathbb{R}^{(q \times d) \times (q \times d) \times (q \times d)} \to \mathbb{R}^{(q \times d)}, ~~
(Q, K, X') ~\mapsto ~ X^* = A(Q, K, X'),
\label{eq:attention}
\end{equation}
```
where $Q \in \mathbb{R}^{q \times d}$ is a matrix of query vectors, $K \in \mathbb{R}^{q \times d}$ is a matrix of key vectors,
 $X' \in \mathbb{R}^{q \times d}$ is a matrix of value vectors and $X^* \in \mathbb{R}^{q \times d}$, 
 the output of the attention mechanism, is a new matrix of values that have been reweighted according to their importance for the modelling problem. 
 As noted above, several different options for the mapping in Equation \ref{eq:attention} exist in the literature; here we describe the scaled dot-product self-attention mechanism 
 used in @Vaswani2017a. In more detail, consider that $X'$ is a matrix of covariates relating to a regression problem, which, 
 in the case of tabular data, will usually be composed of embeddings of the same dimension for several different categorical or numerical variables.
 In the case discussed here we have $q$ covariates mapped to embeddings of dimension $d$. For clarity, note that in Equation \ref{eq:embed_to_nn} 
 we have used the symbol $X'$ to denote the vector of embeddings and numerical inputs which is passed into the neural network. 
 We slightly overload this notation and, in what follows, we now use the symbol $X'$ to refer to the matrix of embeddings of dimension $q \times d$.

We wish to apply the attention mechanism to $X'$ to assign weights to the most important embeddings for the regression problem; 
this is done from the "perspective" of each covariate represented by the $q$ rows of $X'$. To do this, we formulate a so-called query matrix $Q$ 
which contains information about the regression problem at hand; this is usually done by passing $X'$ through a neural network trained to derive 
queries from the covariates. Similarly, the matrix $K$ contains the relevance of each row of $X'$ for the regression problem and 
is also usually derived by passing each row of $X'$ through another neural network. Finally, the matrix $X'$ can be left as is, or processed through another neural network. 
The first step of the attention mechanism is to derive a matrix $A \in \mathbb{R}^{q \times q}$ of attention scores:

```{=latex}
\begin{equation}
A: \mathbb{R}^{(q \times d) \times (q \times d)} \to \mathbb{R}^{q \times q}, ~~
(Q, K) ~\mapsto ~ A = \frac{Q~K^T}{\sqrt{d}},
\label{eq:attention_scores}
\end{equation}
```

and where the division by $\sqrt{d}$ is done element-wise. The attention scores $A$ are then normalized, or made to sum to unity, 
across the columns of $A$ using a so-called softmax function i.e.

@@

```{=latex}
\begin{equation}
X^*: \mathbb{R}^{(q \times q) \times (q \times d)} \to \mathbb{R}^{q \times d}, ~~
(A^*, X') ~\mapsto ~ X^* = A^*~X'.
\label{eq:attention_output}
\end{equation}
```
<!-- In most applications of attention, Equation \ref{eq:attention_scores} is modified by dividing the attention scores by $d^{\frac{1}{2}}$. -->

@@

Self-attention is the main building block of the current state of the art model for natural language processing, which is the Transformer model [@Vaswani2017a]. 
A basic Transformer model applies self-attention to the covariate matrix, $X'$ which is fed into the model, to produce a new matrix of covariates, $X^*$. 
These two matrices, the input matrix $X'$ and the output matrix $X^*$ are added (in an element-wise manner) together and then normalized. 
This matrix is then further processed through a neural network and subsequently, these processed covariates are used in the machine learning task. These stages of the 
Transformer model help to make the optimization of the network easier than applying scaled self-attention directly to the matrix of covariates $X'$.

Multiple Transformer network "blocks" can be added to a network to create a deep Transformer network.

@@

```{=latex}
\begin{equation}
\begin{split}
  X^* &= A(X') \\
  z^1 &= \sigma(a_1.X^* + b_1),
\end{split}
\label{eq:embed_to_attn_to_nn}
\end{equation}
```
where $X' \in \mathbb{R}^{q \times d}$ is the matrix of embeddings after concatenating $q$ embeddings of dimension $d$ together, $X^*$ 
represents the matrix of embeddings $X'$ after processing with attention and $A$ is the attention function defined in Equations \ref{eq:attention} - \ref{eq:attention_output}. 
Below, we demonstrate the effect of adding this simple application of attention to a tabular model.

@Huang2020 refines this approach in a model they call the TabTransformer, which implements two main changes to the approach shown in Equations \ref{eq:embed_to_attn_to_nn}. 
The first of these is that, instead of using the matrix of embeddings after processing with self-attention ($X'$) directly in the model, 
the output of a Transformer model is used instead. The second of these refinements is that for each of the $q$ covariates with an embedding in the matrix $X'$, 
we include a new row embedding which denotes which covariate each embedding relates to. Thus, in the TabTransformer model, we augment 
the embedding matrix with extra columns, so that the dimension of this matrix is $q \times (d + d_{col})$, where $d_{col}$ is the dimension of the row embedding.
@Huang2020 find that including this identifier improves the performance of the model. We compare the results of the TabTransformer model to the other models tested below.

Finally, a different way of incorporating attention into the modeling of tabular data has been proposed by @Ark2019, who refer to their model as TabNet. 
In summary, the main idea of the TabNet model is to try emulate the excellent performance of decision trees on tabular data modelling using deep neural networks. 
This is done by structuring the neural networks to emulate a key aspect of decision tree modelling, which is the efficient selection of relevant covariates
for the model task. Within TabNet, an attention based model is used to select features by estimating attention scores for each covariate within $X$. 
The covariates are multiplied by the attention scores to down-weight less relevant features for the prediction problem. Other less important details of the Tabnwet
model are given in @Ark2019.

# Predictive Modeling with Embeddings {#modelling}

@@

<!-- # Extending the model {#extend}

Investigate approaches for using embedding layers for numeric variables, or numeric variables that were captured as 
categorical variables (Can the modelling of numerical variables benefit from the application of embedding layers?) How 
should this best be done? CatBoost - cut into 256 groups smoothness Map numeric to a dense layer and use that as 
embedding - basis expansion Add a distribution to the embeddings Can insights from credibility theory lead to enhanced embeddings? Take credibility mixture over embeddings Cluster embeddings PCA of embeddings
-->
# Attention based modelling {#attention_models}
